---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sakura

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/sakura)](https://CRAN.R-project.org/package=sakura)
[![R-CMD-check](https://github.com/shikokuchuo/sakura/workflows/R-CMD-check/badge.svg)](https://github.com/shikokuchuo/sakura/actions)
[![Codecov test coverage](https://codecov.io/gh/shikokuchuo/sakura/graph/badge.svg)](https://app.codecov.io/gh/shikokuchuo/sakura)
<!-- badges: end -->

```
  ________  
 /\ sa    \
/  \  ku   \
\  /    ra /
 \/_______/
```

### Extension to R Serialization

Extends the functionality of R serialization by augmenting the built-in reference hook system. This enhanced implementation allows an integrated single-pass operation that combines R serialization with third-party serialization methods.

Facilitates the serialization of even complex R objects, which contain non-system reference objects, such as those accessed via external pointers, to enable their use in parallel and distributed computing.

This package was a request from a meeting of the [R Consortium](https://r-consortium.org/) [Marshalling and Serialization Working Group](https://github.com/RConsortium/marshalling-wg/) held at useR!2024 in Salzburg, Austria. It is designed to eventually provide a common framework for marshalling in R.

It extracts the functionality embedded within the [mirai](https://github.com/shikokuchuo/mirai) async framework for use in other contexts.

### Installation

Install the current release from CRAN:

```{r cran, eval=FALSE}
install.packages("sakura")
```

Or the development version using:

```{r devinstall, eval=FALSE}
pak::pak("shikokuchuo/sakura")
```

### Overview

Some R objects by their nature cannot be serialized, such as those accessed via an external pointer.

Using the [`arrow`](https://arrow.apache.org/docs/r/) package as an example:

```{r arrowfail,error=TRUE}
library(arrow, warn.conflicts = FALSE)
obj <- list(as_arrow_table(iris), as_arrow_table(mtcars))

unserialize(serialize(obj, NULL))
```
In such cases, `sakura::serial_config()` can be used to create custom serialization configurations, specifying functions that hook into R's native serialization mechanism for reference objects ('refhooks').

```{r arrowcfg}
cfg <- sakura::serial_config(
  "ArrowTabular",
  arrow::write_to_raw,
  function(x) arrow::read_ipc_stream(x, as_data_frame = FALSE)
)
```
This configuration can then be supplied as the 'hook' argument for `sakura::serialize()` and `sakura::unserialize()`.
```{r arrowpass}
sakura::unserialize(sakura::serialize(obj, cfg), cfg)
```
This time, the arrow tables are handled seamlessly.

Other types of serialization function are vectorized and in this case, the configuration should be created specifying `vec = TRUE`. Using `torch` as an example:

```{r torchfail, error=TRUE}
library(torch)
x <- list(torch_rand(5L), runif(5L))

unserialize(serialize(x, NULL))
```
Base R serialization above fails, but `sakura` serialization succeeds:

```{r torchpass}
cfg <- sakura::serial_config("torch_tensor", torch::torch_serialize, torch::torch_load, vec = TRUE)

sakura::unserialize(sakura::serialize(x, cfg), cfg)
```

### C Interface

A C level interface is provided. A public header file `sakura.h` is available in `inst/include` for all packages that declare sakura in `LinkingTo`. This may be used in the following way:

```c
#include <sakura.h>

sakura_sfunc sakura_serialize;
sakura_ufunc sakura_unserialize;

// runtime initialization:
sakura_serialize = (sakura_sfunc) R_GetCCallable("sakura", "sakura_serialize");
sakura_unserialize = (sakura_ufunc) R_GetCCallable("sakura", "sakura_unserialize");
```

### Acknowledgements

We would like to thank in particular:

- [R Core](https://www.r-project.org/contributors.html) for providing the interface to the R serialization mechanism.
- [Luke Tierney](https://github.com/ltierney/) and [Mike Cheng](https://github.com/coolbutuseless) for their meticulous efforts in documenting the serialization interface.
- [Daniel Falbel](https://github.com/dfalbel) for discussion around an efficient solution to serialization and transmission of torch tensors.

--

Please note that this project is released with a [Contributor Code of Conduct](https://shikokuchuo.net/sakura/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
